Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-08 paper.tex
updated list stuff

\subsection{List functions}
Our list function domain consists of hand-designed concepts from which we
sample tasks, so there is no ``ground truth'' formal DSL from which these
tasks are generated. Additionally, few of these tasks are concatenative ---
many of them represent distinct operations. We found that this domain was
difficult to tackle without starting from a relatively rich DSL, and even
then it was still hard to solve all tasks. All tasks are solved by functions
that take as input an integer or a list of integers, and have as output
either a Boolean, an integer, a list of Booleans, or a list of integers.
Examples of these tasks are in Table~\ref{listexamples}.

\begin{table}
\centering
\begin{tabular}{| l | l | l |}
  \hline
  \emph{name} & \emph{input} & \emph{output} \\
  \hline
  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  range & 3 & [1\, 2\, 3] \\
  has-2 & [4\, 5\, 7\, 4] & \code{false} \\
  has-head-in-tail & [4\, 5\, 7\, 4] & \code{true} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  \hline
\end{tabular}
\caption{Examples of simple tasks from our list function domain. We used
  15 input/output examples for each task to reduce ambiguity in the task's
  function, ensuring solutions achieve the desired concept.}
\label{listexamples}
\end{table}

@@

\begin{table}[t]
  \centering
  Subset of learned DSL\\
  \begin{tabular}{l}
    \toprule
    $(f_0$ \code{i $\ell$}$) = $\code{(singleton (index i $\ell$))}\\
    \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
    $(f_1$ \code{i $\ell$}$) = $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
    \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
    $(f_2$ \code{n $\ell$}$) = $\code{(any (lambda (x) (= n x)) $\ell$)}\\
    \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
    $(f_3$ \code{i $\ell$}$) = $\code{(index (negate i) (sort $\ell$))}\\
    \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
    $(f_4$ \code{n $\ell$}$) = $\code{(mapi (lambda (i x) (mod x n)) $\ell$)}\\
    $(f_5$ \code{k $\ell$}$) = $\code{(mapi (lambda (i x) (+ x k)) $\ell$)}\\
    $(f_6$ \code{k n $\ell$}$) = $\code{($f_4$ n ($f_5$ k $\ell$)}\\
    \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\
  \bottomrule
  \end{tabular}
  \caption{Some induced primitives from the list DSL}\label{listinduced}
\end{table}

>>>>

\subsection{List functions}
Our list function domain consists of hand-designed concepts from which we
sample tasks, so there is no ``ground truth'' formal DSL from which these
tasks are generated. Additionally, few of these tasks are concatenative ---
many of them represent distinct operations.
Tasks in this domain are solved by functions that take as input an integer
or a list of integers, and have as output either a Boolean, an integer, a
list of Booleans, or a list of integers.  There are 225 tasks in total, some
of which are shown in Table~\ref{listexamples}.

@@

We found that this domain was difficult to tackle without starting from a
relatively rich DSL, and even then it was still hard to solve all tasks. Our
algorithm, like human learners, requires a spectrum of problems ranging from
easy to hard. Without this curriculum, \system cannot get off the ground.
The behavior with the base DSL illustrates this, because the initial set of
primitives is not paired with a sufficiently large and graded curriculum.

\begin{table}
\centering
\begin{tabular}{lll}
  \toprule
  Name & Input & Output \\\midrule
  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  range & 3 & [1\, 2\, 3] \\
  has-2 & [4\, 5\, 7\, 4] & \code{false} \\
  has-head-in-tail & [4\, 5\, 7\, 4] & \code{true} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  \bottomrule
\end{tabular}
\caption{Simple tasks from our list function domain. We used 15 input/output
  examples for each task to reduce ambiguity in the task's function,
  ensuring solutions achieve the desired concept.}
\label{listexamples}
\end{table}

\begin{table}[t]
  \centering
  \begin{tabular}{l}
    \toprule
    $(f_0$ \code{i $\ell$}$) = $\code{(singleton (index i $\ell$))}\\
    \hspace{0.5cm}($f_0$: \emph{Put the $i$-th index into a list})\\
    $(f_1$ \code{i $\ell$}$) = $\code{(++ $\ell$ ($f_0$ i $\ell$))}\\
    \hspace{0.5cm}($f_1$: \emph{Append the $i$-th index})\\
    $(f_2$ \code{n $\ell$}$) = $\code{(any (lambda (x) (= n x)) $\ell$)}\\
    \hspace{0.5cm}($f_1$: \emph{Whether $n$ appears in the list})\\
    $(f_3$ \code{i $\ell$}$) = $\code{(index (negate i) (sort $\ell$))}\\
    \hspace{0.5cm}($f_3$: \emph{Get the $i$-th largest number})\\
    $(f_4$ \code{n $\ell$}$) = $\code{(mapi (lambda (i x) (mod x n)) $\ell$)}\\
    $(f_5$ \code{k $\ell$}$) = $\code{(mapi (lambda (i x) (+ x k)) $\ell$)}\\
    $(f_6$ \code{k n $\ell$}$) = $\code{($f_4$ n ($f_5$ k $\ell$)}\\
    \hspace{0.5cm}($f_6$: \emph{Caesar shift of $k$ in integers modulo $n$})\\
  \bottomrule
  \end{tabular}
  \caption{Some induced primitives for list functions.}\label{listinduced}
\end{table}
